---
title: "Aides Drug"
author: "Kathleen Castro"
date: "2/21/2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.


```{r}
druggies <- filter(triUmp, DrugUser== "yes")
```


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~treatment + infected,data=druggies)
barchartGC(~treatment + infected,data=druggies, type="percent")

```
    
```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")

```
  
According to the first graphs, it depicts that taking Placebo gives a higher percentage of getting aids. When looking at the second graph, it looks like there are more cases of aids when placebo is taken. There appears to be a higher percentage compared to the drug.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=druggies)
rowPerc(table1)
colPerc(table1)
```
  
According to the provided data, placebo carries a higher average number compared to the drug on both results. The first result has the drug at 24.29 and the placebo at 75.71 when looking at aids. The second results shows the drug being at 2.89 while the placebo carries a higher number at 7.77.
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
Looking at the results the Chi-squared data concludes that the P value chance is 0.0000001242. The fishers exact test data also gives us a estimated ratio of 0.35 The overall amount of users that took placebo and obtaining aids is 106 while users who took drugs only had 34.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
In conclusion, we see that users who took Placebo are at a higher risk of getting aids compared to other drugs. Looking at the first graph we see a higher percentage of aids when placed under the placebo mark. This data shows the comparison of percentage between the two types. The second data shows the average number of users who have taken placebo or the drug and which ones got aids from it. The final data chart gives us a ratio we can use that expresses the difference between the two drugs where Placebo carries a higher risk of giving aids to it's users.
